
---
title: "Eyal Berman - Workshop in Data Science"
output: html_notebook
---
##### PLEASE SET THE WORKING FOLDER TO THE FOLDER ALL THE FILES ARE IN BEFORE RUNNING THE SCRIPT#####

```{r}
skip_computation <- TRUE

list.of.packages <-
  c(
  "ggplot2",
  "dplyr",
  "tidyr",
  "randomForest",
  # "caret",
  "hydroGOF",
  # "ETLUtils",
  "reshape"
  )
  new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
  if (length(new.packages)){
    install.packages(new.packages)
  }
  lapply(list.of.packages, require, character.only = TRUE)
  source('ds_project_functions.R')
  
```



## Loading the data tables ##
```{r}
target_data <- read.csv("WorldBank/Millennium_development_goals.csv")
poverty <- read.csv('WorldBank/poverty_and_equity.csv', na.strings = c('..',''))
gender <- read.csv('WorldBank/gender_statistics_female.csv', na.strings = c('..',''))
africa_infra <- read.csv('WorldBank/Africa_Infrastructure.csv', na.strings = c('..',''))
africa_development <- read.csv('WorldBank/africa_development_indicators.csv', na.strings = c('..',''))
health_nutrition <- read.csv('WorldBank/health_and_nutrition.csv', na.strings = c('..',''))
```

## Preparing the tables for training (transpose + fix values) ##
```{r}
target_data <- prepare_target_data(target_data)
poverty <- prepare_data_df_to_pivot(poverty)
gender <- prepare_data_df_to_pivot(gender)
africa_infra <- prepare_data_df_to_pivot(africa_infra)
africa_infra$Country.Name <- gsub(" ","",gsub("_[A-Z]+","", gsub("\\d.","", (africa_infra$Country.Name))))
health_nutrition <- prepare_data_df_to_pivot(health_nutrition)
africa_development <- prepare_african_development_df(africa_development)

africa_countries <- africa_development$Country.Name %>% unique()
```


## Joining the tables into one big dataset, then filter it to include African countries only ##
```{r}
joined_table <- create_joined_table()
african_joined_table <- joined_table %>% filter(Country.Name.x %in% africa_countries)
cat(paste("  The size of the initial data is",nrow(african_joined_table),"rows with",ncol(african_joined_table),                                        "features"))
```

# Data Preperation #


##Remove trivial fields##
Remove fields that directly relate to life expectency and death.
The list was comprised by running the model and looking at the imporant fields list - then identifying the non-interesting features and removing them
```{r}
ColumnsToRemove <- c("Mortality","Death.rate","Life.expectancy","Lifetime.risk","Survival.to","Population.ages","Number.of.under.five.deaths","Number.of.infant.deaths","Number.of.neonatal.deaths")

ColumnsToRemoveNumbers <- unique (grep(paste(ColumnsToRemove,collapse="|"), 
                        colnames(african_joined_table)))
african_joined_table <- african_joined_table[ ,  -ColumnsToRemoveNumbers]
```

##Remove columns with lots of null values##
I've tried several options (0.5, 0.9, 0.1), and found out the removing columns with more than 30% null values
works best
```{r}
african_joined_table_with_null <- african_joined_table
african_joined_table <-remove_large_null_columns(african_joined_table, 0.3)
```


##Prepare data for model training##
Impute null values, removing rows with no observation data
```{r}
african_joined_table$Country.Code <- as.factor(african_joined_table$Country.Code)
african_joined_table.imputed <- na.roughfix(african_joined_table)

suppressWarnings(african_joined_table.imputed$Mort_Rate_under_5 <-       
  as.numeric(paste(african_joined_table.imputed$Mort_Rate_under_5)))
 # Remove NA rows
african_joined_table.imputed <- african_joined_table.imputed %>% filter(Mort_Rate_under_5 >= 0)
cat(paste("The final size of the data is",nrow(african_joined_table),"rows with",ncol(african_joined_table),                                        "features"))
```

## Randomly split the data ##
Split the data into training (80%) and testing (20%) sets
```{r}
smp_size <- floor(0.8 * nrow(african_joined_table.imputed))

set.seed(123)
train_ind <- sample(seq_len(nrow(african_joined_table.imputed)), size = smp_size)
train <- african_joined_table.imputed[train_ind, ]
test <- african_joined_table.imputed[-train_ind, ]
```

## Training the model ##
```{r}
if (skip_computation) {
    rf.model <- readRDS("./final_rf_model.rds")
} else {
  rf.model <-train_model(300,round(sqrt(length(train))))
}

# saveRDS(rf.model, "./final_rf_model.rds")
```
## Evaluating the model ##
The most important features, based on the %incMSE are:
```{r, fig.width = 6}
prediction <- predict(rf.model, test)
compare_df <- data.frame(prediction,test$Mort_Rate_under_5, prediction - test$Mort_Rate_under_5)
varImpPlot(rf.model,type=1)
```
*%incMSE = The increase in MSE the would have happened if we permutated the feature. The larger the number, the more the feature contributes to the accuracy of the table

The following shows the IncNodePurity
```{r, fig.width=6}
varImpPlot(rf.model,type=2)
```
 *IncNodePurity = The average increase in node purity in the tree splits featuring the feature
 
It is clear that although these 2 metrics measures different things, they both find roughly the same features as the most important ones ##

##rMSE##
```{r}
rmse_prediction <- rmse(compare_df$prediction, compare_df$test.Mort_Rate_under_5)
mean_mortality_rate <- mean(african_joined_table.imputed$Mort_Rate_under_5)
median_mortality_rate <- median(african_joined_table.imputed$Mort_Rate_under_5)
cat(" rMSE is",round(rmse_prediction,2),"\n","mean_mortality_rate is ",round(mean_mortality_rate,2),"\n",
    "median_mortality_rate is ",round(median_mortality_rate,2))

```

#Fine Tune the Model Parameters#
```{r}
if (skip_computation) {
  print_model_parameter_testing_results()
} else {
train_predict_and_evaluate_model(test,600,100)
train_predict_and_evaluate_model(test,600,30)
train_predict_and_evaluate_model(test,600,20)
train_predict_and_evaluate_model(test,600,10)
train_predict_and_evaluate_model(test,300,24)
train_predict_and_evaluate_model(test,300,30)
train_predict_and_evaluate_model(test,300,20)
train_predict_and_evaluate_model(test,300,10)
train_predict_and_evaluate_model(test,150,24)
train_predict_and_evaluate_model(test,150,30)
train_predict_and_evaluate_model(test,150,20)
train_predict_and_evaluate_model(test,150,10)
train_predict_and_evaluate_model(test,50,24)
train_predict_and_evaluate_model(test,50,30)
train_predict_and_evaluate_model(test,50,20)
train_predict_and_evaluate_model(test,50,10)
}
```
As we can see, the results don't differ all that much, but choosing the number of features to be around the square root of the total of number of features does present the best results. For example, using 100 as the number of feature consistently perform worse then using 24

# Let's look at some of the main features #
```{r,fig.width=5}
 plot_chart("Adolescent.fertility.rate..births.per.1.000.women.ages.15.19.","fertelity rate 15-19",c(1990,2010))
```




```{r,fig.width=5}
plot_chart("Age.dependency.ratio..young","Age dependency ratio",c(2000))

```

*Age dependency ratio, young, is the ratio of younger dependents--people younger than 15--to the working-age population--those ages 15-64. Data are shown as the proportion of dependents per 100 working-age population.
(from the world bank, http://data.worldbank.org/indicator/SP.POP.DPND.YG)

```{r,fig.width=5}
 plot_chart("Immunization..Pol3....of.one.year.old.children.","Polio Immunization rate among children under 1 year old",c(2000))

```

```{r,fig.width=5}
 plot_chart("Telephone.mainlines..per.1.000.people.","Telephone lines per 1,000 people",c(1990,2010))
```

```{r,fig.width=5}
 plot_chart("People.practicing.open.defecation..urban....of.urban.population.","open defecation",c(1990))
```

*People practicing open defecation refers to the percentage of the population defecating in the open, such as in fields, forest, bushes, open bodies of water, on beaches, in other open spaces or disposed of with solid waste.
(The World Bank, <http://data.worldbank.org/indicator/SH.STA.ODFC.UR.ZS>)

```{r,fig.width=5}ne
 plot_chart("Newborns.protected.against.tetanus....","Newborns.protected.against.tetanus",c(1990,2010))
```
*Newborns protected against tetanus are the percentage of births by women of child-bearing age who are immunized against tetanus.
(The World Bank, <http://data.worldbank.org/indicator/SH.VAC.TTNS.ZS>)

```{r,fig.width=5}
 plot_chart("School.enrollment..primary..female....gross.","School.enrollment.primary.female",c(1990,2010))
```

```{r,fig.width=5}
 plot_chart("Ratio.of.female.to.male.labor.force.participation.rate......modeled.ILO.estimate.","Ratio of female to male labor force participation rate",c(1990,2010))
```
* Labor force participation rate is the proportion of the population ages 15 and older that is economically active: all people who supply labor for the production of goods and services during a specified period.
(The World Bank, <http://data.worldbank.org/indicator/SL.TLF.CACT.FM.ZS>)


